---
output:
  github_document:
    html_preview: false
---
<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r results='asis', echo = FALSE, eval = TRUE}
d <- packageDescription("collections")
```

```{r results="asis", echo = FALSE, eval = TRUE}
cat(c("# ", paste(trimws(strsplit(d$Title, "\n")[[1]]), collapse = " ")))
```

[![Github Action](https://github.com/randy3k/collections/workflows/build/badge.svg?branch=master)](https://github.com/randy3k/collections)
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/collections)](https://cran.r-project.org/package=collections)
[![codecov](https://codecov.io/gh/randy3k/collections/branch/master/graph/badge.svg)](https://codecov.io/gh/randy3k/collections)
[![](http://cranlogs.r-pkg.org/badges/grand-total/collections)](https://cran.r-project.org/package=collections)

Documentation: [http://randy3k.github.io/collections](https://randy3k.github.io/collections)

```{r results="asis", echo = FALSE, eval = TRUE}
cat(d$Description)
```

## Related R packages

[dequer](https://CRAN.R-project.org/package=dequer) implements double
ended queues and it supports arbitrary R objects. However, it uses
`R_PreserveObject` and `R_ReleaseObject` heavily which could be an issue
for long queues.

[datastructures](https://CRAN.R-project.org/package=datastructures) uses
‘Boost’ and ‘STL’ data types to implement queues and hashmaps. For some
reasons, it is often slow as shown in the benchmark.

[liqueueR](https://CRAN.R-project.org/package=liqueueR) implements
queues in pure R code.

[hash](https://CRAN.R-project.org/package=hash) uses R environment to 
create hash tables. As indicated by the author of 
[fastmap](https://CRAN.R-project.org/package=fastmap),
a small amount of memory may be leaked.

[hashmap](https://CRAN.R-project.org/package=hashmap) provides
hashmap for atomic vectors using boost library.

[fastmap](https://CRAN.R-project.org/package=fastmap) provides hashmap
which does not leak memory. The current implementation of collection::Dict
is actually inspired by it. However, a more
efficient hash table library [tommy](https://github.com/amadvance/tommyds/)
is used.
