---
output:
  github_document:
    html_preview: false
---
<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r results='asis', echo = FALSE, eval = TRUE}
d <- packageDescription("collections")
```

```{r results="asis", echo = FALSE, eval = TRUE}
cat(c("# ", paste(trimws(strsplit(d$Title, "\n")[[1]]), collapse = " ")))
```

[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/collections)](https://cran.r-project.org/package=collections)

Github: [https://github.com/randy3k/collections](https://github.com/randy3k/collections)

Documentation: [http://randy3k.github.io/collections](https://randy3k.github.io/collections)

```{r results="asis", echo = FALSE, eval = TRUE}
cat(d$Description)
```

## Related R packages

[dequer](https://CRAN.R-project.org/package=dequer) implements double ended queues and it supports arbitary R objects. However, it uses `R_PreserveObject` and `R_ReleaseObject` heavily which could be an issue for long queues.

[datastructures](https://CRAN.R-project.org/package=datastructures) uses 'Boost' and 'STL' data types to implement queues and hashmaps.

[liqueueR](https://CRAN.R-project.org/package=liqueueR) implements queues using R lists.

[hash](https://CRAN.R-project.org/package=hash) uses  `new.env( hash = TRUE)` to create hash tables.

[hashmap](https://CRAN.R-project.org/package=hashmap) provides fast hashmap for atomic vectors.
