---
title: "Benchmarking 'collections'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking 'collections'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of this document is to compare the performances of some of the existing packages.

- [dequer](https://CRAN.R-project.org/package=dequer) `r packageDescription("dequer")$Version`implements double
    ended queues and it supports arbitrary R objects. However, it uses
    `R_PreserveObject` and `R_ReleaseObject` heavily which could be an issue
    for long queues.    
- [datastructures](https://CRAN.R-project.org/package=datastructures )
    `r packageDescription("datastructures")$Version` uses
    ‘Boost’ and ‘STL’ data types to implement queues and hashmaps. For some
    reasons, it is often slow as shown in the benchmark.
- [liqueueR](https://CRAN.R-project.org/package=liqueueR)
    `r packageDescription("liqueueR")$Version` implements
    queues in pure R code which explains its slowness.
- [hash](https://CRAN.R-project.org/package=hash)
    `r packageDescription("hash")$Version` uses R environments to 
    create hash tables. It is [known](https://r-lib.github.io/fastmap/#memory-leak-examples) that regular R enviroments leak memory.
- [hashmap](https://CRAN.R-project.org/package=hashmap)
    `r packageDescription("hashmap")$Version` provides
    hashmap for atomic vectors using boost library.
- [fastmap](https://CRAN.R-project.org/package=fastmap)
    `r packageDescription("fastmap")$Version`. The current implementation of collection::dict
    is actually inspired by it. However, a more
    efficient hash table library [tommy](https://github.com/amadvance/tommyds/)
    is used.
- We also compare to serveral base R implementations based on lists or environments, see [here](https://github.com/randy3k/collections/blob/master/vignettes/benchmark.Rmd) for details.


```{r global_options, include=FALSE}
requireNamespace("bench")
library(tidyverse)
options(digits = 3)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


## Queue

```{r, echo = FALSE}
queue <- function() {
    self <- environment()
    q <- NULL
    n <- NULL
    push <- function(item) {
        if (is.null(item)) {
            q[n + 1] <<- list(item)
        } else {
            q[[n + 1]] <<- item
        }
        n <<- n + 1
        invisible(self)
    }
    pop <- function() {
        if (n == 0) stop("queue is empty")
        v <- q[[1]]
        q <<- q[-1]
        n <<- n - 1
        v
    }
    peek <- function() {
        if (n == 0) stop("queue is empty")
        q[[1]]
    }
    clear <- function() {
        q <<- list()
        n <<- 0
        self
    }
    clear()
    self
}
```

```{r queue, echo = FALSE, cache = TRUE}
bench_queue <- bench::press(
    n = c(10, 50, 100, 200, 500),
    bench::mark(
        `dequer::queue` = {
            q <- dequer::queue()
            pushback <- dequer::pushback
            pop <- dequer::pop
            x <- rnorm(n)
            for (i in 1:n) pushback(q, x[i])
            for (i in 1:n) pop(q)
        },
        `datastructures::queue` = {
            q <- datastructures::queue()
            insert <- datastructures::insert
            pop <- datastructures::pop
            x <- rnorm(n)
            for (i in 1:n) insert(q, x[i])
            for (i in 1:n) pop(q)
        },
        `liqueueR::Queue` = {
            q <- liqueueR::Queue$new()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:n) q$pop()
        },
        `base::list` = {
            q <- queue()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:n) q$pop()
        },
        `collections::queue` = {
            q <- collections::queue()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:n) q$pop()
        },
        check = FALSE
    )) %>%
    mutate(expression = fct_reorder(
            as.character(expression), median, .fun = mean, .desc = TRUE))
```

```{r, echo = FALSE}
bench_queue %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("push and pop n times") + ylab("time")
bench_queue %>%
    filter(expression %in% c("base::list", "dequer::queue", "collections::queue")) %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("(zoom in)") + ylab("time")
```

## Stack

```{r, echo = FALSE}
stack <- function() {
    self <- environment()
    q <- NULL
    n <- NULL

    push <- function(item) {
        if (is.null(item)) {
            q[n + 1] <<- list(item)
        } else {
            q[[n + 1]] <<- item
        }
        n <<- n + 1
        invisible(self)
    }
    pop <- function() {
        if (n == 0) stop("stack is empty")
        v <- q[[n]]
        q <<- q[-n]
        n <<- n - 1
        v
    }
    clear <- function() {
        q <<- list()
        n <<- 0
        invisible(self)
    }
    clear()
    self
}
```

```{r stack, echo = FALSE, cache = TRUE}
bench_stack <- bench::press(
    n = c(10, 50, 100, 200, 500),
    bench::mark(
        `dequer::stack` = {
            q <- dequer::stack()
            push <- dequer::push
            pop <- dequer::pop
            x <- rnorm(n)
            for (i in 1:n) push(q, x[i])
            for (i in 1:n) pop(q)
        },
        `datastructures::stack` = {
            q <- datastructures::stack()
            insert <- datastructures::insert
            pop <- datastructures::pop
            x <- rnorm(n)
            for (i in 1:n) insert(q, x[i])
            for (i in 1:n) pop(q)
        },
        `liqueueR::Stack` = {
            q <- liqueueR::Stack$new()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:n) q$pop()
        },
        `base::list` = {
            q <- stack()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:n) q$pop()
        },
        `collections::stack` = {
            q <- collections::stack()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:n) q$pop()
        },
        check = FALSE
    )) %>%
    mutate(expression = fct_reorder(
            as.character(expression), median, .fun = mean, .desc = TRUE))
```

```{r, echo = FALSE}
bench_stack %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("push and pop n times") + ylab("time")
bench_stack %>%
    filter(expression %in% c("base::list", "dequer::stack", "collections::stack")) %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("(zoom in)") + ylab("time")
```


## Deque

```{r, echo = FALSE}
deque <- function() {
    self <- environment()
    q <- NULL
    n <- NULL

    push <- function(item) {
        if (is.null(item)) {
            q[n + 1] <<- list(item)
        } else {
            q[[n + 1]] <<- item
        }
        n <<- self$n + 1
        invisible(self)
    }
    pushleft <- function(item) {
        q <<- c(list(item), q)
        n <<- n + 1
        invisible(self)
    }
    pop <- function() {
        if (n == 0) stop("deque is empty")
        v <- q[[n]]
        q <<- q[-n]
        n <<- n - 1
        v
    }
    popleft <- function() {
        if (n == 0) stop("deque is empty")
        v <- q[[1]]
        q <<- q[-1]
        n <<- n - 1
        v
    }
    clear <- function() {
        q <<- list()
        n <<- 0
        invisible(self)
    }
    clear()
    self
}
```

```{r deque, echo = FALSE, cache = TRUE}
bench::press(
    n = c(10, 50, 100, 200, 500),
    bench::mark(
        `dequer::deque` = {
            q <- dequer::deque()
            push <- dequer::push
            pop <- dequer::pop
            popback <- dequer::popback
            x <- rnorm(n)
            for (i in 1:n) push(q, x[i])
            for (i in 1:floor(n / 2)) {
                pop(q)
                popback(q)
            }
        },
        `base::list` = {
            q <- deque()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:floor(n / 2)) {
                q$pop()
                q$popleft()
            }
        },
        `collections::deque` = {
            q <- collections::deque()
            x <- rnorm(n)
            for (i in 1:n) q$push(x[i])
            for (i in 1:floor(n / 2)) {
                q$pop()
                q$popleft()
            }
        },
        check = FALSE
    )) %>%
    mutate(expression = fct_reorder(
            as.character(expression), median, .fun = mean, .desc = TRUE)) %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("push and pop n times") + ylab("time")
```


## Priority Queue

```{r priority_queue, echo = FALSE, cache = TRUE}
bench::press(
    n = c(10, 50, 100, 200, 500),
    bench::mark(
        `liqueueR::PriorityQueue` = {
            q <- liqueueR::PriorityQueue$new()
            x <- rnorm(n)
            p <- sample.int(n, n, replace = TRUE)
            for (i in 1:n) q$push(x[i], p[i])
            for (i in 1:n) q$pop()
        },
        `datastructures::binomial_heap` = {
            q <- datastructures::binomial_heap(key.class = "integer")
            insert <- datastructures::insert
            pop <- datastructures::pop
            x <- rnorm(n)
            p <- sample.int(n, n, replace = TRUE)
            for (i in 1:n) insert(q, p[i], x[i])
            for (i in 1:n) pop(q)
        },
        `collections::priority_queue` = {
            q <- collections::priority_queue()
            x <- rnorm(n)
            p <- sample.int(n, n, replace = TRUE)
            for (i in 1:n) q$push(x[i], p[i])
            for (i in 1:n) q$pop()
        },
        check = FALSE
    )) %>%
    mutate(expression = fct_reorder(
            as.character(expression), median, .fun = mean, .desc = TRUE)) %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("push and pop n times") + ylab("time")
```


## Dictionary

```{r dict, echo = FALSE, cache = TRUE}
bench_dict <- bench::press(
    n = c(10, 50, 100, 200, 500),
    bench::mark(
        `datastructures::hashmap` = {
            h <- datastructures::hashmap()
            insert <- datastructures::insert
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) insert(h, k[i], x[i])
            for (i in sample(1:n)) h[k[i]]
        },
        `hash::hash` = {
            h <- hash::hash()
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h[[k[i]]] <- x[i]
            for (i in sample(1:n)) h[[k[i]]]
        },
        `hashmap::hashmap` = {
            h <- hashmap::hashmap(keys = character(0), values = numeric(0))
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h[[k[i]]] <- x[i]
            for (i in sample(1:n)) h[[k[i]]]
        },
        `fastmap::fastmap` = {
            h <- fastmap::fastmap()
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h$set(k[i], x[i])
            for (i in sample(1:n)) h$get(k[i])
        },
        `base::new.env` = {
            h <- new.env(parent = emptyenv())
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h[[k[i]]] <- x[i]
            for (i in sample(1:n)) get0(k[i], envir = x, inherits = FALSE)
        },
        `collections::dict` = {
            h <- collections::dict()
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h$set(k[i], x[i])
            for (i in sample(1:n)) h$get(k[i])
        },
        check = FALSE
    )) %>%
    mutate(expression = fct_reorder(
            as.character(expression), median, .fun = mean, .desc = TRUE))
```

```{r, echo = FALSE}
bench_dict %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("push and pop n times") + ylab("time")
bench_dict %>%
    filter(expression %in% c(
        "hash::hash", "fastmap::fastmap", "collections::dict", "base::new.env")) %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("(zoom in)") + ylab("time")
```

Note that `base::new.env` suffers from [memory leak](https://r-lib.github.io/fastmap/#memory-leak-examples) issue.


## Ordered dict

```{r ordered_dict, echo = FALSE, cache = TRUE}
bench::press(
    n = c(10, 50, 100, 200, 500),
    bench::mark(
        `base::list` = {
            h <- list()
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h[[k[i]]] <- x[i]
            for (i in sample(1:n)) h[[k[i]]]
        },
        `collections::ordered_dict` = {
            h <- collections::ordered_dict()
            x <- rnorm(n)
            k <- as.character(seq_len(n))
            for (i in 1:n) h$set(k[i], x[i])
            for (i in sample(1:n)) h$get(k[i])
        },
        check = FALSE
    )) %>%
    mutate(expression = fct_reorder(
            as.character(expression), median, .fun = mean, .desc = TRUE)) %>%
    ggplot(aes(x = n, y = median)) +
    geom_line(aes(color = expression)) +
    scale_colour_brewer(palette = "Set2", direction = -1) +
    ggtitle("set and get n times") + ylab("time")
```

Note that `ordered_dict` grows linearly in `n` but but `list` does not.