---
title: "Benchmarking 'collections'"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarking 'collections'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of this document is to compare the performances of some of the existing packages.
See [benchmark.Rmd](https://github.com/randy3k/collections/blob/master/vignettes/benchmark.Rmd) for the details.

We are comparing these packages with `collections`.

- [dequer](https://CRAN.R-project.org/package=dequer) 
    `r packageDescription("dequer")$Version`
- [datastructures ](https://CRAN.R-project.org/package=datastructures ) 
    `r packageDescription("datastructures")$Version`
- [liqueueR](https://CRAN.R-project.org/package=liqueueR) 
    `r packageDescription("liqueueR")$Version`
- [hash](https://CRAN.R-project.org/package=hash)
    `r packageDescription("hash")$Version`
- [hashmap](https://CRAN.R-project.org/package=hashmap)
    `r packageDescription("hashmap")$Version`


```{r, echo = FALSE}
library(microbenchmark)
options(digits = 3)
```

## Queue

```{r, echo = FALSE}
test_dequer_queue_push_pop <- function(n) {
    q <- dequer::queue()
    pushback <- dequer::pushback
    pop <- dequer::pop
    for (i in 1:n) pushback(q, rnorm(1))
    for (i in 1:n) pop(q)
    q
}
test_datastructures_queue_push_pop <- function(n) {
    q <- datastructures::queue("numeric")
    insert <- datastructures::insert
    pop <- datastructures::pop
    for (i in 1:n) insert(q, rnorm(1))
    for (i in 1:n) pop(q)
    q
}
test_liqueueR_queue_push_pop <- function(n) {
    q <- liqueueR::Queue$new()
    for (i in 1:n) q$push(rnorm(1))
    for (i in 1:n) q$pop()
    q
}
test_collections_queue_push_pop <- function(n) {
    q <- collections::Queue$new()
    for (i in 1:n) q$push(rnorm(1))
    for (i in 1:n) q$pop()
    q
}
```

```{r}
microbenchmark(
    dequer = test_dequer_queue_push_pop(2000),
    datastructures = test_datastructures_queue_push_pop(2000),
    liqueueR = test_liqueueR_queue_push_pop(2000),
    collections = test_collections_queue_push_pop(2000),
    times = 10,
    unit = "relative"
)
```

## Stack


```{r, echo = FALSE}
test_dequer_stack_push_pop <- function(n) {
    q <- dequer::queue()
    pushback <- dequer::pushback
    pop <- dequer::pop
    for (i in 1:n) pushback(q, rnorm(1))
    for (i in 1:n) pop(q)
    q
}
test_datastructures_stack_push_pop <- function(n) {
    q <- datastructures::stack("numeric")
    insert <- datastructures::insert
    pop <- datastructures::pop
    for (i in 1:n) insert(q, rnorm(1))
    for (i in 1:n) pop(q)
    q
}
test_liqueueR_stack_push_pop <- function(n) {
    q <- liqueueR::Stack$new()
    for (i in 1:n) q$push(rnorm(1))
    for (i in 1:n) q$pop()
    q
}
test_collections_stack_push_pop <- function(n) {
    q <- collections::Stack$new()
    for (i in 1:n) q$push(rnorm(1))
    for (i in 1:n) q$pop()
    q
}
```

```{r}
microbenchmark(
    dequer = test_dequer_stack_push_pop(2000),
    datastructures = test_datastructures_stack_push_pop(2000),
    liqueueR = test_liqueueR_stack_push_pop(2000),
    collections = test_collections_stack_push_pop(2000),
    times = 10,
    unit = "relative"
)
```

## Deque

```{r, echo = FALSE}
test_dequer_deque_push_pop <- function(n) {
    q <- dequer::deque()
    push <- dequer::push
    pop <- dequer::pop
    popback <- dequer::popback
    for (i in 1:n) push(q, rnorm(1))
    for (i in 1:floor(n / 2)) {
        pop(q)
        popback(q)
    }
    q
}
test_collections_deque_push_pop <- function(n) {
    q <- collections::Deque$new()
    for (i in 1:n) q$push(rnorm(1))
    for (i in 1:floor(n / 2)) {
        q$pop()
        q$popleft()
    }
    q
}
```

```{r}
microbenchmark(
    dequer = test_dequer_deque_push_pop(5000),
    collections = test_collections_deque_push_pop(5000),
    times = 10,
    unit = "relative"
)
```

## Dict / hashmap

```{r, echo = FALSE}
test_datastructures_dict_set_get <- function(n) {
    h <- datastructures::hashmap("character", "numeric")
    insert <- datastructures::insert
    get <- datastructures::get
    for (i in 1:n) insert(h, as.character(i), rnorm(1))
    for (i in 1:n) get(h, as.character(i))
    h
}
test_hash_dict_set_get <- function(n) {
    h <- hash::hash()
    for (i in 1:n) h[[as.character(i)]] <- rnorm(1)
    for (i in 1:n) h[[as.character(i)]]
    h
}
test_hashmap_dict_set_get <- function(n) {
    h <- hashmap::hashmap(keys = character(0), values = numeric(0))
    for (i in 1:n) h[[as.character(i)]] <- rnorm(1)
    for (i in 1:n) h[[as.character(i)]]
    h
}
test_collections_dict_set_get <- function(n) {
    h <- collections::Dict$new()
    for (i in 1:n) h$set(as.character(i), rnorm(1))
    for (i in 1:n) h$get(as.character(i))
    h
}
```

```{r}
microbenchmark(
    datastructures = test_datastructures_dict_set_get(5000),
    hash = test_hash_dict_set_get(5000),
    hashmap = test_hashmap_dict_set_get(5000),
    collections = test_collections_dict_set_get(5000),
    times = 10,
    unit = "relative"
)
```
